Light versus Dark in \cb{Strong-Lens Galaxies}: Dark matter halos that are rounder than their stars

### 2015-01-20 shapes/paper.tex
Included remarks by co-authors

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{myaasmacros}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{multirow}

@@

%Code definition:
\def\Glass{{\sc Glass}}
\def\PixeLens{{\sc PixeLens}}

\title[Light versus Dark in Strong Lensing Galaxies]{Light versus Dark in Strong Lensing Galaxies: Dark matter halos that are rounder than their stars}

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{1,2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France
}

@@

\begin{abstract}
We measure the shape and alignment of stars and dark matter in 11 strong
lensing galaxies, finding in all cases that the dark matter halos are rounder
than the light distribution over the range $R_e < R < 5R_e$.  Averaging over
larger radii, the lenses become increasingly elliptical both in their dark
matter and stellar components, as defined using the ratio $s$ of the largest to
smallest eigenvalues of the 2D moment of inertia tensor of the projected
lensing mass.  While dark matter halos are never more elliptical than $s_{dm} =
1.15$, their stars can extend to $s_* > 1.4$ Three systems, in particular, have
very high stellar ellipticity ($s_* > 1.6$) and correspondingly high alignment
between light and dark. One of these -- B1608 -- is a known merging pair; we
suggest that the other two (B0712 and B2016) may also be recent post-merger
systems.  Galaxies with high dark matter ellipticity and weak external shear
show very strong alignment between light and dark; those with strong shear
($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since
isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must
explain the origin of both very round dark matter halos and highly misaligned
strong lensing systems. Such misalignments also present a new challenge for alternative
gravity theories in which the light and dark must necessarily be highly
correlated.

@@

\begin{keywords}
Gravitational lensing: strong --- galaxies: structure
\end{keywords}

@@

The ellipticity and shape of stars relative to their host dark matter halo encodes information both about our cosmological model ($\Lambda$CDM) and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter halos that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces halos that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If the visible light is the only source of gravity in a galaxy, then we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can at least in principle be broken.

Strong lensing provides a unique probe of the alignment and shape of the total mass distribution in galaxies \citep[e.g.][]{1986ApJ...310..568B,1992grle.book.....S,1998ApJ...509..561K,2000ApJ...543..131K,2006ApJ...649..599K,2007AJ....134..668A,2008MNRAS.383..857F,2010ApJ...724..511A,2012MNRAS.424..104L}. For `red and dead' ellipticals that are largely devoid of gas, their baryonic content can be mapped through stellar population synthesis modelling of their light distribution alone \citep[e.g.][]{2005ApJ...623L...5F,2006ApJ...640..662T,2008MNRAS.383..857F}. Furthermore, such systems are dense enough to produce strong lensing, opening up the possibility of directly comparing the light and mass in these galaxies \citep{1998ApJ...509..561K,2008MNRAS.383..857F,2009ApJ...690..670T,2012A&A...538A..99S}. Previous work in the literature has found that the light and mass are well-aligned (though a mis-match of up to 10$^\circ$ is not uncommon; e.g. \citealt{2012A&A...538A..99S}). However, results on the ellipticity of light and mass agree less well, with \citet{2012A&A...538A..99S} finding a strong correlation and \citet{1998ApJ...509..561K} and \citet{2008MNRAS.383..857F} finding none. It is difficult, however, to compare the results between these different studies because they use different lens modelling techniques; different definitions of ellipticity; and different radii over which the shapes and alignments are probed. Furthermore, none to date have applied their methodology to mock data to determine the robustness of the results.

Weak lensing can also be used to probe the shape and alignment between light and dark, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter halos appear to be rounder then their light distributions, with some weak evidence for alignment. Both measurements are challenging, however, and more recent data appear to be at odds with this early work, favouring dark matter halos that are more elliptical than their stars \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}.

Recently, we introduced a new non-parametric lens tool, \Glass\ \citep{2014arXiv1401.7990C}. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in strong lensing systems if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass contributes significantly to the potential. In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter}  in these lensing galaxies, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stars often dominate the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between light and dark is made possible by the fact that \Glass\ uses the light distribution as a prior on the mass map, ensuring that the dark matter mass is always positive.

This paper is organised as follows. In \S\ref{sec:glass}, we briefly review the \Glass\ code. In \S\ref{sec:data}, we present our data compilation with references. In \S\ref{sec:results}, we present our results. Finally, in \S\ref{sec:conclusions} we discuss the implications of these results and we present our main conclusions.

@@

\Glass{} is a new `non-parametric' gravitational lens modelling framework
\citep{2014arXiv1401.7990C} generates a distribution of
mass models given a set of basic assumptions, or priors. Hence, \Glass\ does not yield the most probable mass model. Rather, it calculates an ensemble of all allowed models given the data and assumed priors. In this context,
`non-parametric' really just mean that many more parameters than data constraints are used
such that system of equations describing the lensing mass distribution is
under-constrained. \Glass{} shares some aspects with an earlier code
\PixeLens{} \citep{2004AJ....127.2604S,2008ApJ...679...17C}, but has a completely
different code base. Its key improvements upon \PixeLens{} which are relevant
for this paper include:

\begin{enumerate}
\item A new uniform sampling algorithm for high
    dimensional spaces \citep{2012MNRAS.425.3077L}. This allows for large
    ensembles of $>10,000$ models to be efficiently generated. 
%\item \Glass\ provides a modular framework that allows new priors to be added
%    and modified easily.
%\item The basis functions approximating a model can be easily changed (in this
%    paper, we assume pixels as in \PixeLens). 
\item With so many models in the final ensemble, we can afford to apply
    non-linear constraints (for example stellar kinematic data; or the removal of
    models with spurious extra images) to accept/reject models in a post-processing
    step.
\item The central region of the mass map can have a higher resolution to more
    efficiently capture steep models.
\item Stellar density can be used as an additional lower-bound constraint on the models. 
%\item Point or extended mass objects can be placed in the field.
\end{enumerate}

%\Glass{} is described in detail in the code paper: \citet{2014arXiv1401.7990C}.
In this paper, the default priors used to model each lens are listed below.  Table
\ref{tab:lenspriors} identifies lenses for which we have modified these defaults.
For each lens we assume:

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map as lower mass bounds in every pixel.
\item The local density gradient is always pointing within 50$^{\circ}$ to the center and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item The fixed cosmology $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1]$.
\end{enumerate}

@@

\begin{enumerate}
\item[(vii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.

% \item[(viii)] External point masses to model satellite galaxies outside but
% close to the Einstein radius.

\item[(viii)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$ where 
the surface mass profile can be assumed to be symmetric (3+1 quad
configurations and the double \textit{Q0957}).

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
We employ the same shape definition as in \citep{2014arXiv1401.7990C}. The shape parameter $s$ is defined as the ratio of the two eigenvalues $\lambda_1$ and $\lambda_2$, i.e.
\begin{equation}\label{eq:shapeestimate}
    s \equiv \frac{\lambda_{1}}{\lambda_{2}}, \ \ \ \mathrm{where} \ \lambda_{1} \geq \lambda_{2},
\end{equation}
of the inertia tensor of a mass distribution $M(\boldsymbol{\theta})$
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of the half-light radius $R_e$. $R_e$ is the Petrosian radius containing half of the flux of the total light distribution. As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii -- e.g. effects such as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} can cause differences in the shape estimates between the central and the outer parts of the system.

Throughout this paper, we consider only the masses in the pixels up to one pixel length past the outermost image. We generate in each case an ensemble of 10,000 mass models.

\section{Data}\label{sec:data}
Stellar mass maps are taken from \citep{2011ApJ...740...97L}. In that work, the light of the galaxy images in all available HST bands was first fitted using using GALFIT \citep{2002AJ....124..266P}, and the resulting light maps were used to obtain stellar mass maps using a synthesis of the stellar-population models of \citep{2003MNRAS.344.1000B}. The population synthesis marginalised over the star formation epoch and time scales and over the stellar metallicity. The IMF was taken to be the log-normal form given by \citep{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the inferred shape of the stellar mass.

As shown in \citep{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561} out of the sample presented in \citep{2011ApJ...740...97L}. This leads to a sample of 11 lenses. We discuss each of these systems, next.

\textit{Q0047-2808} (hereafter \textit{0047}) is a luminous early-type galaxy \citep{1996MNRAS.278..139W}. It is part a group of 9 members, all of which are spectroscopically confirmed \citep{2011ApJ...726...84W}.

\textit{MG0414+0534} (hereafter \textit{0414}) is a passively evolving early-type galaxy \citep{1999AJ....117.2034T}. A very close luminous satellite galaxy was found north-west of the lens \citep{1993AJ....105....1S}. It, however, is more likely a foreground object \citep{2011MNRAS.413L..86C}.

\textit{B0712+472} (hereafter \textit{0712}) is an early-type galaxy \citep{1998AJ....115..377F}.

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. X-ray emission can be detected and an analysis of which yields a temperature of 2.3 keV. There is also a satellite galaxy to the northwest \citep{2000ApJ...544L..35K}.

\textit{Q0957+561} (hereafter \textit{0957}) is a cD galaxy lying close to the centre of a cluster with a high spiral galaxy-fraction \citep[e.g.][]{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}. Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the centre of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the centre of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay \citep[e.g.][]{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, a disagreement at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

\textit{PG1115+080} (hereafter \textit{1115}) is an early-type galaxy \citep{2005ApJ...626...51Y}. It has measured time delays. In this work use recent estimates of the time delays by \cite{2010MNRAS.406.2764T} that differ from previously found values by e.g. \cite{1997ApJ...489...21B}. The environment of the lens was analysed thoroughly \citep{2006ApJ...641..169M,2011ApJ...726...84W}. It is part of a small group of 13 members. Also, X-ray emission was detected from the corresponding group and it yields a temperature of 0.8 keV \citep{2004ApJ...610..686G}.

\textit{B1422+231} (hereafter \textit{1422}) is an early-type galaxy \citep{1996ApJ...462L..53I}. Although there are measured time delays \citep{2001MNRAS.326.1403P}, they seem to deviate significantly from theoretical expectations \citep{2003AJ....126...29R}. We choose therefore not to include them in the analysis. The lens is part of a group with 16 spectroscopically confirmed member galaxies \citep{2006ApJ...641..169M}. Using newer data, an additional member could be found \citep{2011ApJ...726...84W}. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

\textit{B1608+656} (hereafter \textit{1608}) consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analysed and a group with 8 (9 if merging galaxies are counted individually) was found \citep{2006ApJ...642...30F}. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seem to indicate four other groups along the line of sight.

\textit{MG2016+112} (hereafter \textit{2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster that consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The cluster shows a high density of galaxies close to the lens in a south-east direction.

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{myaasmacros}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{color}

@@

%Code definition:
\def\Glass{{\sc Glass}}
\def\PixeLens{{\sc PixeLens}}
\newcommand{\cb}[1]{{\color{red} \textbf{#1}}}

\title[Light versus Dark in \cb{Strong-Lens Galaxies}]{Light versus Dark in \cb{Strong-Lens Galaxies}: Dark matter halos that are rounder than their stars}
% "Light versus dark matter in strong-lens galaxies: Dark matter halos may be rounder than their stellar distributions"

\author[Bruderer et al.]
{\parbox{\textwidth}{Claudio Bruderer,$^{1}$\thanks{E-mail: \texttt{claudio.bruderer@phys.ethz.ch}}
Justin I. Read,$^{1,2}$
Prasenjit Saha,$^{3}$
Jonathan Coles,$^{4}$
Emilio E. Falco,$^{5}$
Ignacio Ferreras$^{6}$ and
Dominik Leier$^{7}$}\vspace{0.4cm}\\
\parbox{\textwidth}{$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France\\
$^{5}$Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA 02138, USA\\
$^{6}$Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, Surrey RH5 6NT, UK\\
$^{7}$Dipartimento di Fisica e Astronomia, Università di Bologna, viale Berti Pichat 6/2, 40127, Bologna, Italy}}

@@

\begin{abstract}
We measure the shape and alignment of the stellar and dark matter mass distribution in 11 \cb{strong-lens galaxies}, finding in all cases that the dark matter halos are rounder than the stellar mass distribution over the range $R_e < R < 5R_e$.  Averaging over larger radii, the lenses become increasingly elliptical both in their dark matter and stellar components, as defined \cb{using the ratio of the smallest to largest eigenvalues} of the 2D moment of inertia tensor of the projected \cb{lens mass}.  While dark matter halos are never more elliptical \cb{than $e_{dm} = 0.3$}, their stars can extend \cb{to $e_* > 0.3$}. Three systems, in particular, \cb{have a high stellar ellipticity ($e_* > 0.4$)} and correspondingly high alignment between luminous and dark matter distribution. One of these -- B1608 -- is a known merging pair; we suggest that the other two (B0712 and B2016) may also be recent post-merger systems.  Galaxies with high dark matter ellipticity and weak external shear show very strong alignment between light and dark; those with strong shear ($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must explain the origin of both very round dark matter halos and highly misaligned \cb{strong-lens systems}. Such misalignments also present a new challenge for alternative gravity theories in which the stellar and dark matter components must necessarily be highly correlated.

@@

\begin{keywords}
Gravitational lensing: strong --- galaxies: structure --- galaxies: haloes --- galaxies: formation --- galaxies: elliptical and lenticular, cD
\end{keywords}

@@

The ellipticity and shape of the stellar component relative to their host dark matter halo encodes information both about our cosmological model ($\Lambda$CDM) and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter halos that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces halos that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If the visible light is the only source of gravity in a galaxy, then we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can at least in principle be broken.

Strong lensing provides a unique probe of the alignment and shape of the total mass distribution in galaxies \citep[e.g.][]{1986ApJ...310..568B,1992grle.book.....S,1998ApJ...509..561K,2000ApJ...543..131K,2006ApJ...649..599K,2007AJ....134..668A,2008MNRAS.383..857F,2010ApJ...724..511A,2012MNRAS.424..104L}. For `red and dead' ellipticals that are largely devoid of gas, their baryonic content can be mapped through stellar population synthesis modelling of their light distribution alone \citep[e.g.][]{2005ApJ...623L...5F,2006ApJ...640..662T,2008MNRAS.383..857F}. Furthermore, such systems are dense enough to produce strong lensing \cb{effects}, opening up the possibility of directly comparing the light and mass in these galaxies \citep{1998ApJ...509..561K,2008MNRAS.383..857F,2009ApJ...690..670T,2012A&A...538A..99S}. Previous work in the literature has found that the light and mass are well-aligned (though a mis-match of up to 10$^\circ$ is not uncommon; e.g. \citealt{2012A&A...538A..99S}). However, results on the ellipticity of light and mass agree less well, with \citet{2012A&A...538A..99S} finding a strong correlation and \citet{1998ApJ...509..561K} and \citet{2008MNRAS.383..857F} finding none. It is difficult, however, to compare the results between these different studies because they use different lens modelling techniques; different definitions of ellipticity; and different radii over which the shapes and alignments are probed. Furthermore, none to date have applied their methodology to mock data to determine the robustness of the results.

Weak lensing can also be used to probe the shape and alignment between the luminous and dark matter distributions, but only for `stacked' galaxies \citep{2000astro.ph..6281B,2000ApJ...538L.113N}. \citet{2004ApJ...606...67H} applied this idea to data from the Red-Sequence Cluster Survey (RCS) to measure the first weak lensing signal of halo flattening. They found that dark matter halos appear to be rounder than the stellar mass distribution, with some weak evidence for alignment. Both measurements are challenging however, and more recent data appear to be at odds with this early work, favouring dark matter halos that are more elliptical than their stellar mass component \citep{2006MNRAS.370.1008M,2007ApJ...669...21P,2012A&A...545A..71V}.

Recently, we introduced a new non-parametric lens tool, \Glass\ \citep{2014arXiv1401.7990C}. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in \cb{lens systems} if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass contributes significantly to the potential. In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter} in these \cb{lens galaxies}, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stellar component often dominates the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between the stellar and dark matter components is made possible by the fact that \Glass\ uses the stellar mass distribution as a prior on the mass map, ensuring that the dark matter mass is always positive.

This paper is organised as follows. In \S\ref{sec:glass}, we briefly review the \Glass\ code \cb{and define our method to assess shape and alignment of lens galaxies}. In \S\ref{sec:data}, we present our data compilation with references. In \S\ref{sec:results}, we present our results. Finally, in \S\ref{sec:conclusions} we discuss the implications of these results and we present our main conclusions.

@@

\Glass\ is a new `non-parametric' gravitational lens modelling framework \citep{2014arXiv1401.7990C} that generates a distribution of mass models given a set of basic assumptions, or priors. Hence, \Glass\ does not yield the most probable mass model. Rather, it calculates an ensemble of all allowed models given the data and assumed priors. In this context, `non-parametric' really just means that many more parameters than data constraints are used such that the system of equations describing the \cb{lensing mass} distribution is under-constrained. \Glass\ shares some aspects with an earlier code \PixeLens\ \citep{2004AJ....127.2604S,2008ApJ...679...17C}, but has a completely different code base. Its key improvements upon \PixeLens{} which are relevant for this paper include:

\begin{enumerate}
\item A new uniform sampling algorithm for high dimensional spaces \citep{2012MNRAS.425.3077L}. This allows for large ensembles of $>10,000$ models to be efficiently generated. 
\item With so many models in the final ensemble, we can afford to apply non-linear constraints (for example stellar kinematic data; or the removal of models with spurious extra images) to accept/reject models in a post-processing step.
\item The central region of the mass map can have a higher resolution to more efficiently capture steep models.
\item Stellar density can be used as an additional lower-bound constraint on the models. 
\end{enumerate}

In this paper, the default priors used to model each lens are listed below.  Table
\ref{tab:lenspriors} identifies lenses for which we have modified these defaults.
For each lens we assume:

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map as lower mass bounds in every pixel.
\item The local density gradient \cb{always points within 50$^{\circ}$ of the center of the distribution,} and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item The fixed cosmology $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1]$.
\item \cb{The pixel radius is set to be 12 pixels.}
\end{enumerate}

@@

\begin{enumerate}
\item[(viii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.
\item[(ix)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$, where the surface mass profile can be assumed to be symmetric ((3+1)-quad configurations and the double \textit{Q0957}).

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
\cb{We employ a shape definition related to the definition used by \cite{2014arXiv1401.7990C}. The ellipticity parameter $e$ is defined as} %the ratio of the two eigenvalues $\lambda_2$ and $\lambda_1$, i.e.
\begin{equation}\label{eq:shapeestimate}
    e \equiv 1 - \frac{\lambda_{2}}{\lambda_{1}}, \ \ \ \mathrm{where} \ \lambda_{1} \geq \lambda_{2},
\end{equation}
of the inertia tensor of a mass distribution $M(\boldsymbol{\theta})$
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of the half-light radius $R_e$. \cb{$R_e$ is the half-light radius measured with respect to the Petrosian flux.} As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii -- e.g. effects such as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} can cause differences in the shape estimates between the central and the outer parts of the system.

Throughout this paper, we consider only the masses in the pixels up to one pixel length past the outermost image. We generate in each case an ensemble of 10,000 mass models. \cb{Each of these models is analyzed individually, yielding a distribution of constrained quantities.}

\section{Data}\label{sec:data}
Stellar mass maps are taken from \cite{2011ApJ...740...97L}. In that work, the \cb{surface brightness distribution} of the galaxy images in all available HST bands was first fitted using using GALFIT v2.03b \citep{2002AJ....124..266P}, and the resulting \cb{flux} maps were used to obtain stellar mass maps using the synthesis of the stellar-population models of \cite{2003MNRAS.344.1000B}. The population synthesis is marginalised over the star formation epoch and time scales and over the stellar metallicity. The IMF was taken to be the log-normal form given by \cite{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} \cb{but not the shape of the inferred stellar mass distribution \citep[unless the IMF presents significant intrinsic deviations locally, see e.g.][]{2015MNRAS.447.1033M}. We note that the effect of recent claims towards a non-universal IMF on stellar M/L is still under debate. While dynamical studies seem to point at contributions at the level of the Salpeter IMF or slightly heavier \citep{2013MNRAS.432.1862C}, spectral line strength constraints can still accommodate a wider range of the stellar M/L depending on the functional form of the IMF \citep{2013MNRAS.429L..15F}.}

\cb{The original sample we consider, is presented in \cite{2011ApJ...740...97L} and consists of 21 lens systems. It is a subsample of the CfA-Arizona Space Telescope LEns Survey\footnote{http://www.cfa.harvard.edu/castles/} (CASTLES) sample. As shown in \citep{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561}. This leads to a sample of 11 lenses. We discuss each of these systems, next.}

\textit{Q0047-2808} (hereafter \textit{0047}) is a luminous early-type galaxy \citep{1996MNRAS.278..139W}. It is part of a group of 9 members, all of which are spectroscopically confirmed \citep{2011ApJ...726...84W}.

\textit{MG0414+0534} (hereafter \textit{0414}) is a passively evolving early-type galaxy \citep{1999AJ....117.2034T}. A very close luminous satellite galaxy was found north-west of the lens \citep{1993AJ....105....1S}. It, however, is more likely a foreground object \cb{with a redshift of ~0.38 \citep{2011MNRAS.413L..86C}. This is sufficiently distant to still contribute to the lensing effect.}

\textit{B0712+472} (hereafter \textit{0712}) is an early-type galaxy \cb{\citep{1998MNRAS.296..483J,1998AJ....115..377F}. There is a foreground group of about 10 member galaxies at $z\sim0.3$, which could affect the modelling \citep{2002AJ....123..627F}.}

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost \cb{spherical} early-type galaxy \cb{\citep{1997A&A...317L..13B,2012A&A...538A..99S}.} It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. X-ray emission can be detected and an analysis of which yields a temperature of 2.3 keV. There is also a satellite galaxy to the northwest \citep{2000ApJ...544L..35K}.

\textit{Q0957+561} (hereafter \textit{0957}) is a cD galaxy lying close to the centre of a cluster with a high spiral galaxy-fraction \citep[e.g.][]{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}. Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the centre of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg. These results are consistent with the centre of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay \citep[e.g.][]{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, a disagreement at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

\textit{PG1115+080} (hereafter \textit{1115}) is an early-type galaxy \cb{\citep{1980Natur.285..641W,2005ApJ...626...51Y}.} It has measured time delays \cb{\citep[see e.g.][]{1997ApJ...475L..85S}.} In this work\cb{, we} use recent estimates of the time delays by \cite{2010MNRAS.406.2764T} that differ from previously found values by e.g. \cite{1997ApJ...489...21B}. The environment of the lens was analysed thoroughly \citep{2006ApJ...641..169M,2011ApJ...726...84W}. It is part of a small group of 13 members. Also, X-ray emission was detected from the corresponding group and it yields a temperature of 0.8 keV \citep{2004ApJ...610..686G}.

\textit{B1422+231} (hereafter \textit{1422}) is an early-type galaxy \cb{\citep{1992MNRAS.259P...1P,1994AJ....107...28Y}.} Although there are measured time delays \citep{2001MNRAS.326.1403P}, they seem to deviate significantly from theoretical expectations \citep{2003AJ....126...29R}. We choose therefore not to include them in the analysis. The lens is part of a group with 16 spectroscopically confirmed member galaxies \citep{2006ApJ...641..169M}. Using newer data, \cb{additional members} could be found \citep{2011ApJ...726...84W}. \cite{2004ApJ...610..686G} also detect X-ray emission from the corresponding group at a temperature of 1.0 keV.

\textit{B1608+656} (hereafter \textit{1608}) \cb{was first reported by \cite{1995ApJ...447L...5M}. It} consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analysed and a group with 8 (9 if merging galaxies are counted individually) was found \citep{2006ApJ...642...30F}. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seem to indicate four other groups along the line of sight.
